Weighted Model Counting with Conditional Probabilities

### 2020-07-04 paper.tex
File format conversion (and updated data analysis)
2004-pgm dataset

@@

\paragraph{Experimental setup.}
\begin{itemize}
\item
  Algorithms\footnote{\url{http://beyondnp.org/pages/solvers/model-counters-exact/}}
  (highlighted the initial setup, can expand later):
  \begin{itemize}
  \item ADDMC \cite{DBLP:conf/aaai/DudekPV20} (rediscovered the
    multiplicativity of BAs in different words) (with optimal settings)
  \item Cachet \cite{DBLP:conf/sat/SangBBKP04}
  \item c2d \cite{DBLP:conf/ecai/Darwiche04}
  \item d4 \cite{DBLP:conf/ijcai/LagniezM17} (closed source, boo!)
  \item miniC2D  \cite{DBLP:conf/ijcai/OztokD15}
  \end{itemize}
\item Encodings:
  \begin{itemize}
  \item \texttt{-d02} \cite{DBLP:conf/kr/Darwiche02}
  \item \texttt{-sbk05} \cite{DBLP:conf/aaai/SangBK05}
  \item \texttt{-cd05} \cite{DBLP:conf/ijcai/ChaviraD05}
  \item \texttt{-cd06} \cite{DBLP:conf/sat/ChaviraD06} (supposed to be the best)
  \item mine
  \end{itemize}
  All except mine are from Ace
  3.0\footnote{\url{http://reasoning.cs.ucla.edu/ace/}} and should be compiled
  with \texttt{-encodeOnly} (i.e., don't compile the CNF into an AC) and
  \texttt{-noEclause} (i.e., only use standard syntax) flags.
\item Datasets\footnote{There might be more at
    \url{https://meelgroup.github.io/}.}
  \begin{itemize}
  \item binary Bayesian networks from Sang et
    al.\footnote{\url{https://www.cs.rochester.edu/u/kautz/Cachet/}}
    \cite{DBLP:conf/aaai/SangBK05}
    \begin{itemize}
    \item Grid networks: computing marginal probability of the last node in
      the grid (DNE).
    \item Plan recognition problems (DNE + inst).
    \item Deterministic quick medical reference (DNE + inst).
    \end{itemize}
    For the last two, the task can be to calculate the marginal probability of a
    randomly chosen node (so, repeat at least ten times).
  \item Bayesian networks available with Ace (some binary, some non-binary)
    \begin{itemize}
    \item \texttt{2004-pgm}: NET + inst (\cite{DBLP:journals/ijar/ChaviraDJ06})
    \item \texttt{2005-ijcai}: XDSL/NET/Hugin + inst (\cite{DBLP:conf/ijcai/ChaviraD05})
    \item \texttt{2006-ijar}: NET + inst (\cite{DBLP:journals/ijar/ChaviraDJ06})
    \end{itemize}
  \item ProbLog \cite{DBLP:conf/uai/FierensBTGR11}
  \item probabilistic programs \cite{DBLP:journals/corr/abs-2005-09089}
  \end{itemize}
\end{itemize}

@@

\paragraph{Notes.}
\begin{itemize}
\item Thesis: many important computational problems are solved by encoding them
  as WMC. But if the weight function is extended to `conditional probabilities',
  the problem becomes easier because current approaches need workarounds in
  order to be independent.
\item Shorter thesis: many important problems are encoded as WMC, but they can
  be encoded in a better way if we allow for conditional probabilities.
\item We extend the Gaifman graph to add edges when two variables occur in the
  same CPT (e.g., including the edge from $A$ to $B$ when the CPT is $\Pr(A \mid
  B)$).
\item Extra benefit: one does not need to come up with a way to turn some probability
distribution to into a fully independent one.
\item Trivial CPTs such as $\Pr(A \mid B) = \Pr(\neg A \mid B) = 0.5$, the ADDs
  of which simplify to a single number, are put in the first cluster.
\item Observation: by inspecting the BN, we could identify CPTs that could be
  completely ignored, but maybe good heuristics take care of that anyway.
\item Observation: ADDs have a restrict command. I'm not going to use it, but it
  could be nice.
\item Alternative scenario to consider for future work: \#\SAT{} solver
  generates solutions which are then used to restrict the ADDs that calculate
  probabilities.
\item Important future work: replacing ADDs with
  AADDs\footnote{\url{https://github.com/ssanner/dd-inference}} is likely to
  bring performance benefits. Other extensions:
  \begin{itemize}
  \item FOADDs can represent first order statements;
  \item XADDs can replace WMI for continuous variables;
  \item ADDs with intervals can do approximations.
  \end{itemize}
\item In preliminary results, the SBK05 encoding gives the wrong asnwer in 60
  QMR instances (all with evidence files). Sometimes that answer is $\gg 1$.
\item Similarly, cd06 is supposed to be the fastest, but seems to be one of the
  slowest (along with cd05). Perhaps the best encoding for arithmetic circuits
  has very little to do with the best encoding for CNF WMC solvers?
\item When the Bayesian network has an evidence file, we compute the probability
  of evidence. Otherwise, we pick the last (i.e., last to be mentioned) node in
  the Bayesian network to be the goal. This applies to the Grid data set (as
  intended) and also to two instances of Plan Reconstruction. Check if these two
  answers match.
\end{itemize}

>>>>

\item Are they called Boolean Bayesian networks or binary Bayesian networks?

@@

\paragraph{Experimental setup.}
\begin{itemize}
\item
  Algorithms\footnote{\url{http://beyondnp.org/pages/solvers/model-counters-exact/}}
  \begin{itemize}
  \item ADDMC \cite{DBLP:conf/aaai/DudekPV20} (rediscovered the
    multiplicativity of BAs in different words) (with optimal settings)
  \item Cachet \cite{DBLP:conf/sat/SangBBKP04}
  \item c2d \cite{DBLP:conf/ecai/Darwiche04}
  \item d4 \cite{DBLP:conf/ijcai/LagniezM17} (closed source, boo!)
  \item miniC2D  \cite{DBLP:conf/ijcai/OztokD15}
  \end{itemize}
\item Encodings:
  \begin{itemize}
  \item \texttt{-d02} \cite{DBLP:conf/kr/Darwiche02}
  \item \texttt{-sbk05} \cite{DBLP:conf/aaai/SangBK05}
  \item \texttt{-cd05} \cite{DBLP:conf/ijcai/ChaviraD05}
  \item \texttt{-cd06} \cite{DBLP:conf/sat/ChaviraD06} (supposed to be the best)
  \item mine
  \end{itemize}
  All except mine are from Ace
  3.0\footnote{\url{http://reasoning.cs.ucla.edu/ace/}} and should be compiled
  with \texttt{-encodeOnly} (i.e., don't compile the CNF into an AC) and
  \texttt{-noEclause} (i.e., only use standard syntax) flags.
\item Datasets\footnote{There might be more at
    \url{https://meelgroup.github.io/}.}
  \begin{itemize}
  \item binary Bayesian networks from Sang et
    al.\footnote{\url{https://www.cs.rochester.edu/u/kautz/Cachet/}}
    \cite{DBLP:conf/aaai/SangBK05}
    \begin{itemize}
    \item Grid (networks) (ratio 75 means that 75\% of the nodes are
      deterministic),
    \item Plan recognition (problems),
    \item Deterministic quick medical reference (what do the numbers mean? the
      README doesn't say).
    \end{itemize}
  \item Bayesian networks available with Ace
    \begin{itemize}
    \item \texttt{2004-pgm} \cite{DBLP:journals/ijar/ChaviraDJ06}
    \item \texttt{2005-ijcai} \cite{DBLP:conf/ijcai/ChaviraD05}. The Genie/Smile
      files have their own citation data that I should probably extract. This is
      the only dataset that has some non-binary networks.
    \item \texttt{2006-ijar} \cite{DBLP:journals/ijar/ChaviraDJ06}
    \end{itemize}
  \item ProbLog \cite{DBLP:conf/uai/FierensBTGR11}
  \item probabilistic programs \cite{DBLP:journals/corr/abs-2005-09089}
  \end{itemize}
\end{itemize}

@@

\paragraph{Notes.}
\begin{itemize}
\item Thesis: many important computational problems are solved by encoding them
  as WMC. But if the weight function is extended to `conditional probabilities',
  the problem becomes easier because current approaches need workarounds in
  order to be independent.
\item Shorter thesis: many important problems are encoded as WMC, but they can
  be encoded in a better way if we allow for conditional probabilities.
\item We extend the Gaifman graph to add edges when two variables occur in the
  same CPT (e.g., including the edge from $A$ to $B$ when the CPT is $\Pr(A \mid
  B)$).
\item Extra benefit: one does not need to come up with a way to turn some probability
distribution to into a fully independent one.
\item Trivial CPTs such as $\Pr(A \mid B) = \Pr(\neg A \mid B) = 0.5$, the ADDs
  of which simplify to a single number, are put in the first cluster.
\item Observation: by inspecting the BN, we could identify CPTs that could be
  completely ignored, but maybe good heuristics take care of that anyway.
\item Observation: ADDs have a restrict command. I'm not going to use it, but it
  could be nice.
\item Alternative scenario to consider for future work: \#\SAT{} solver
  generates solutions which are then used to restrict the ADDs that calculate
  probabilities.
\item Important future work: replacing ADDs with
  AADDs\footnote{\url{https://github.com/ssanner/dd-inference}} is likely to
  bring performance benefits. Other extensions:
  \begin{itemize}
  \item FOADDs can represent first order statements;
  \item XADDs can replace WMI for continuous variables;
  \item ADDs with intervals can do approximations.
  \end{itemize}
\item In preliminary results, the SBK05 encoding gives the wrong asnwer in 60
  QMR instances (all with evidence files). Sometimes that answer is $\gg 1$.
\item Similarly, cd06 is supposed to be the fastest, but seems to be one of the
  slowest (along with cd05). Perhaps the best encoding for arithmetic circuits
  has very little to do with the best encoding for CNF WMC solvers?
\item When the Bayesian network has an evidence file, we compute the probability
  of evidence. Otherwise, we pick the last (i.e., last to be mentioned) node in
  the Bayesian network to be the goal. This applies to the Grid data set (as
  intended) and also to two instances of Plan Reconstruction and roughly half of
  2004-PGM.
\end{itemize}
